The 2014 SMT Competition

### 2016-01-11 2014/report/SMTCOMP.tex
Further second-round comments addressed.

In the succeeding sections we describe 
the competition goals (\S\ref{sec:goals}), 
the SMT-LIB language that is the basis for the competition (\S\ref{sec:context}), 
the competition benchmarks (\S\ref{sec:benchmarks}), 
participants (\S\ref{sec:participants}), 
procedure (\S\ref{sec:procedure}),
computational infrastructure (\S\ref{sec:starexec}), 
comparisons with other competitions (\S\ref{sec:OtherCompetitions}), 
the results (\S\ref{sec:results}), 
the place of SMT-COMP in the FLoC Olympic Games (\S\ref{sec:floc}), 
and post-competition activities (\S\ref{sec:post}). 
Section \ref{sec:conclusions} presents observations on this competition and recommendations for the future.

@@

The SMT Competition is a competition among SMT solvers on a set of benchmark logic problems. Each benchmark problem is a combination of definitions and logical assertions expressed with respect to an underlying logical theory and, perhaps, some constraints on the kinds of expressions in that theory. For example,  the logic of linear arithmetic includes the multiplication operation, but only allows multiplication by constants. Each problem is a set of closed formulas over a set of constant or function symbols, possibly including quantification; a solution to the problem is an assignment of each constant and function evaluation to values in a way that satisfies all the problem's assertions. That is, the task is to find a \emph{satisfying assignment} for the benchmark problem or to determine that there is no such assignment. Since the presence of quantified expressions introduces incompleteness, solvers may also produce a potential solution that may be marked as possibly incomplete.

As stated, the goal of SMT-COMP is similar to the goal of the SAT competition. The SMT logic extends SAT by incorporating defined theories, such as the theory of arrays, or of uninterpreted functions, or of arithmetic, or of bit-vectors. In addition, there may be constraints on the set of expressions allowed, such as only linear arithmetic, or only integer difference arithmetic. The theories also define \emph{sorts}, which make the theories a kind of typed first-order logic. Examples of sorts used in current theories are Boolean, Int, Real, bit-vectors of various lengths, and arrays with arbitrary sorts as index and value. Each combination of underlying theories and language constraints is a \emph{logic}.
The names of the logics as used in SMT-LIB are cryptic combinations of initials. For example, AUFLIA is the logic with a
combination of Arrays (A), Uninterpreted Functions (UF), and linear integer arithmetic (LIA). Table \ref{logicAbbreviations} can be used to interpret these names.
\begin{table}
\begin{center}
\begin{tabular}{|l|l|}
\hline
QF\_ & Quantifier-Free \\
A & Arrays \\
UF & Unidentified Functions \\
BV & Bit-Vector \\
L & Linear (arithmetic) \\
N & Non-Linear (arithmetic) \\
IA/RA/IRA & Integer/Real/Mixed arithmetic \\
IDL/RDL & Integer/Real Difference Logic \\
\hline
\end{tabular}
\end{center}
\caption{Abbreviations used in logic names}
\label{logicAbbreviations}
\end{table}

@@

\begin{minipage}{\textwidth}
\begin{quote}
\lstset{frameround=fttt}
\begin{lstlisting}[frame=trBL]
(set-logic UFLIA)
(declare-fun max (Int Int) Int)
(assert (forall ((x Int) (y Int))
   (let ((m (max x y)))
      (and (>= m x) (>= m y) (or (= m x) (= m y))))))
(assert (not (forall ((x Int) (y Int))
   (let ((m (max x y)))
      (= (max (max x y) x) (max x y))))))
(check-sat)
\end{lstlisting}
\end{quote}
\end{minipage}

@@

\begin{figure}
\begin{center}
\begin{tabular}{cc}
\includegraphics[width=.47\textwidth]{smtlib2-file-count-by-size.png}
&
\includegraphics[width=.47\textwidth]{smtlib2-app-file-count-by-size.png}
\\
\\
Non-incremental benchmarks & Application benchmarks
\end{tabular}
\end{center}
\caption{Distribution of SMT-LIB benchmarks by file size.
\label{fig:smtlib-bysize}}
\end{figure}

@@

The numbers of unknown, trivial, and remaining eligible benchmarks are shown in Table \ref{Table:benchmarks}. Note that these numbers vary widely from division to division.

\begin{table}
\caption{Numbers of main-track benchmarks. The second column shows the number of competitive solvers and, in square brackets, the number of demonstration-only solvers. Entries marked * exclude some benchmarks containing partial functions.}
\label{Table:benchmarks}
\centering
\begin{tabular}{|l|r|r|r|r|r|}
\hline
       & \multicolumn{1}{|c|}{\# of} & \multicolumn{4}{|c|}{\# of benchmarks} \\
 Logic & solvers & eligible & unknown  & trivial & total  \\
\hline
ALIA & 3+[1] & 29 & 0 & 13 & 42 \\
AUFLIA & 3+[1] & 4 & 0 & 0 & 4\\
AUFLIRA & 3+[1] & 10791 & 168 & 9055 & 20014 \\
AUFNIRA & 2+[2] & 564 & 468 & 463 & 1495 \\
BV & 2+[1] & 0 & 191 & 0 & 191 \\
LIA & 3+[1] & 46 & 0 & 0 & 46 \\
LRA & 3+[1] & 171 & 450 & 0 & 621\\
NIA & 2+[1] & 9 & 0 & 0 & 9\\
NRA & 2+[1] & 3747 & 66 & 0 & 3813 \\
QF\_ABV & 7+[2] & 6457* & 4191 & 4423 & 15091 \\
QF\_ALIA & 3+[2] & 97 & 0 & 29 & 126 \\
QF\_AUFBV & 2+[2] & 37 & 0 & 0 & 37 \\
QF\_AUFLIA & 4+[2] & 610 & 0 & 399 & 1009\\
QF\_AX & 3+[2] & 335 & 0 & 216 & 551 \\
QF\_BV & 8+[3] & 2488* & 28138 & 546 & 32500 \\
QF\_IDL & 3+[1] & 1315 & 537 & 337 & 2189 \\
QF\_LIA & 4+[3] & 4381 & 1279 & 481 & 6141 \\
QF\_LRA & 4+[2] & 1343 & 208 & 131 & 1682\\
QF\_NIA & 3+[1] & 8327 & 927 & 105 & 9359 \\
QF\_NRA & 3+[1] & 10121 & 1392 & 27 & 11540 \\
QF\_RDL & 3+[1] & 132 & 85 & 38 & 255 \\
QF\_UF & 5+[2] & 4124 & 4 & 2522 & 6650 \\
QF\_UFBV & 2+[2] & 31 & 0 & 0 & 31 \\
QF\_UFIDL & 3+[1] & 311 & 0 & 130 & 441\\
QF\_UFLIA & 4+[2] & 484 & 0 & 114 & 598 \\
QF\_UFLRA & 4+[2] & 1176 & 87 & 367 & 1630 \\
QF\_UFNIA & 2+[1] & 7 & 0 &0 & 7 \\
QF\_UFNRA & 2+[1] & 32 & 11 & 0 & 43\\
UF & 3+[1] & 2830 & 2911 & 7 & 5748 \\
UFBV & 2+[1] & 0 & 191 & 0 & 191 \\
UFIDL & 2+[1] & 49 & 12 & 19 & 80 \\
UFLIA & 3+[1] & 5766 & 5499 & 873 & 12138 \\
UFLRA & 3+[1] & 25 & 0 & 0 & 25\\
UFNIA & 2+[1] & 1587 & 1052 & 712 & 3351 \\
\hline
Total & 18+[3] & 67426 & 47867 & 21007 & 137648 \\
\hline
\end{tabular}
\end{table}

@@

The language includes commands that allow a fine-grained interaction
with the solver, whereby client tools may incrementally push and pop
symbol declarations and assertions while running various
satisfiability checks, inspecting models, or obtaining unsatisfiability
proofs. Benchmarks problems that have more than one
\lstinline{check-sat} command are called \emph{application} or \emph{incremental
  benchmarks}. Not all SMT solvers support all of these interaction
facilities, and the main track of the competition does not use such
application benchmarks.

@@

Benchmarks are also labeled by \emph{category}: simple checks, randomly generated problems from some template (e.g., N-queens problems for various values of N), problems crafted to test a certain capability, and problems obtained from industrial applications. The selection rules favor industrial benchmarks.

@@

The scrambling had another effect, previously unappreciated. Although most of the solvers are deterministic for a given input, scrambling can change the order of search. For example, iterating over the contents of a hashed set may occur in a different order if identifiers have different names. Consequently, the success of the solver on a given benchmark, the time it takes to find an answer, and even whether a latent bug in the solver is triggered may all depend on just how the benchmark is scrambled. Thus the scrambling adds a measure of uncertainty to the competition and prevents full testing on the benchmark set before the competition. Indeed, one solver encountered a bug during the competition that was directly related to the specific, randomly-chosen scrambling used during the competition. The effect of the possible variations in performance caused by scrambling is expected to be balanced by the large of number of benchmarks employed in the competition to compare the solvers, as also observed in the SAT community~\cite{leberre+03}. Nevertheless, this point is worthy of future further investigation.

@@

\begin{table}
\centering\begin{tabular}{|l|lll|}
\hline
Logic & Fraction solved fastest & & \\ \hline
ALIA & 51\% [Z3] & 48\% veriT  & \\
AUFLIA & 66\% [Z3] & 33\% CVC3  & \\
AUFLIRA & 60\% [Z3] & 37\% veriT & 1\% CVC3 \\
AUFNIRA & 72\% [Z3] & 23\% CVC3 & 2\% CVC4 \\
LIA & 43\% CVC4 & 30\% [Z3] & 17\% CVC3 \\
LRA & 42\% CVC4 & 38\% veriT & 18\% [Z3] \\
NIA & 66\% [Z3] & 33\% CVC4  & \\
NRA & 96\% CVC4 & 3\% [Z3]  & \\
QF\_ABV & 78\% Yices2 & 9\% SONOLAR & 7\% Kleaver-STP \\
QF\_ALIA & 95\% Yices2 & 4\% [Z3] & \\
QF\_AUFBV & 81\% Yices2 & 10\% [MathSAT] & 5\% [Z3] \\
QF\_AUFLIA & 60\% Yices2 & 38\% [Z3] & 1\% CVC4 \\
QF\_AX & 92\% Yices2 & 6\% [Z3] & 0\% CVC4 \\
QF\_BV & 33\% Yices2 & 25\% [Z3] & 13\% 4Simp \\
QF\_IDL & 87\% Yices2 & 11\% [Z3] & 0\% veriT \\
QF\_LIA & 76\% Yices2 & 11\% [Z3] & 7\% CVC4 \\
QF\_LRA & 86\% Yices2 & 3\% veriT & 3\% CVC4 \\
QF\_NIA & 90\% [Z3] & 7\% AProVE & 1\% CVC3 \\
QF\_NRA & 88\% [Z3] & 6\% CVC4 & 4\% CVC3 \\
QF\_RDL & 71\% Yices2 & 25\% [Z3] & 3\% veriT \\
QF\_UF & 96\% Yices2 & 1\% veriT & 1\% [Z3] \\
QF\_UFBV & 96\% Yices2 & 3\% [MathSAT]  & \\
QF\_UFIDL & 59\% [Z3] & 36\% Yices2 & 3\% veriT \\
QF\_UFLIA & 97\% Yices2 & 2\% [Z3] & 0\% CVC4 \\
QF\_UFLRA & 91\% Yices2 & 4\% [Z3] & 2\% veriT \\
QF\_UFNIA & 100\% CVC4 & & \\
QF\_UFNRA & 59\% CVC3 & 37\% [Z3] & 3\% CVC4 \\
UF & 37\% veriT & 32\% CVC4 & 24\% [Z3] \\
UFIDL & 57\% CVC4 & 42\% [Z3]  & \\
UFLIA & 56\% [Z3] & 35\% veriT & 7\% CVC4 \\
UFLRA & 95\% veriT & 4\% [Z3]  & \\
UFNIA & 80\% [Z3] & 10\% CVC3 & 9\% CVC4 \\

\hline
\end{tabular}
\caption{Fraction of benchmarks solved fastest (cpu-time) by each solver (top 3)}
\label{tab:speed}
\end{table}

@@

\begin{table}
\centering\begin{tabular}{|l|lll|}
\hline
Logic & Unique contribution & & \\
\hline
ALIA & 0.4253 [Z3] & 0.4080 CVC4 & 0.1667 veriT \\
AUFLIA & 0.3889 CVC4 & 0.2778 [Z3] & 0.2222 CVC3 \\
AUFLIRA & 0.2615 [Z3] & 0.2521 CVC4 & 0.2503 CVC3 \\
AUFNIRA & 0.2509 [Z3] & 0.2506 CVC4 & 0.2497 CVC3 \\
LIA & 0.3949 [Z3] & 0.3514 CVC4 & 0.2319 CVC3 \\
LRA & 0.3070 CVC4 & 0.3070 [Z3] & 0.2895 CVC3 \\
NIA & 0.8333 [Z3] & 0.1667 CVC4 & 0.0000 Yices2 \\
NRA & 0.5071 [Z3] & 0.4929 CVC4 & 0.0000 Yices2 \\
QF\_ABV & 0.1147 Yices2 & 0.1146 Boolector-just... & 0.1144 Boolector-dual... \\
QF\_ALIA & 0.2235 Yices2 & 0.2235 SMTInterpol & 0.2235 [Z3] \\
QF\_AUFBV & 0.2635 [MathSAT] & 0.2635 CVC4 & 0.2500 Yices2 \\
QF\_AUFLIA & 0.1996 Yices2 & 0.1996 [MathSAT] & 0.1996 CVC4 \\
QF\_AX & 0.2000 Yices2 & 0.2000 [MathSAT] & 0.2000 CVC4 \\
QF\_BV & 0.1051 Boolector & 0.0976 STP-Crypto... & 0.0934 abziz\_min\_f... \\
QF\_IDL & 0.2893 [Z3] & 0.2731 Yices2 & 0.2395 CVC4 \\
QF\_LIA & 0.1620 [CVC4-...fix] & 0.1610 [Z3] & 0.1610 SMTInterpol \\
QF\_LRA & 0.1772 CVC4 & 0.1689 Yices2 & 0.1678 SMTInterpol \\
QF\_NIA & 0.4990 [Z3] & 0.4912 AProVE & 0.0084 CVC3 \\
QF\_NRA & 0.7714 [Z3] & 0.1347 CVC3 & 0.0898 CVC4 \\
QF\_RDL & 0.2734 Yices2 & 0.2431 veriT & 0.2431 [Z3] \\
QF\_UF & 0.1451 Yices2 & 0.1449 veriT & 0.1446 CVC4 \\
QF\_UFBV & 0.4140 Yices2 & 0.3172 [Z3] & 0.1720 [MathSAT] \\
QF\_UFIDL & 0.2708 Yices2 & 0.2708 [Z3] & 0.2309 CVC4 \\
QF\_UFLIA & 0.1698 Yices2 & 0.1698 [MathSAT] & 0.1698 CVC4 \\
QF\_UFLRA & 0.1715 [Z3] & 0.1681 Yices2 & 0.1681 [MathSAT] \\
QF\_UFNIA & 0.5000 CVC4 & 0.5000 [Z3] & 0.0000 Yices2 \\
QF\_UFNRA & 0.5802 [Z3] & 0.2099 CVC4 & 0.2099 CVC3 \\
UF & 0.4921 CVC4 & 0.1988 [Z3] & 0.1759 CVC3 \\
UFIDL & 0.5204 [Z3] & 0.4796 CVC4 & 0.0000 Yices2 \\
UFLIA & 0.4069 CVC4 & 0.3502 [Z3] & 0.2425 veriT \\
UFLRA & 0.2857 [Z3] & 0.2381 CVC4 & 0.2381 veriT \\
UFNIA & 0.4156 [Z3] & 0.3860 CVC4 & 0.1984 CVC3 \\
\hline
\end{tabular}
\caption{Contribution from each solver (top 3)}
\label{tab:cont}
\end{table}

@@

\begin{figure}[H]
\begin{center}
\includegraphics[width=.95\textwidth]{QF_BV-timeout-logscale2.png}
\end{center}
\caption{Solver success for QF\_BV benchmarks over time.}
\label{Fig:timeouts}
\end{figure}

@@

The data in Table~\ref{Fig:winners} also shows that most solvers have a similar trend in success as the timeout is increased. For example, most solvers solve about 50\% (ranging from 40\% to 62\%) of the benchmarks that solver solves within the timeout by 1.0 second and
about 75\% (ranging from 65\% to 79\%) of those solved within the timeout by 10 seconds. However, some solvers, such as Boolector, are outliers: Boolector is slow off the mark - solving only 33\% and 61\% of the benchmarks it eventually solves
by 1 second and 10 seconds, respectively, even though by 2400 seconds it solves more benchmarks than any other tool, in this division.

@@

After the competition, David Cok (competition chair) and Clark Barrett and Morgan Deters (SMT-LIB coordinators) collaborated, with the assistance of Aaron Stump (StarExec lead), in attempting to discover the status of the SMT-LIB benchmarks that were marked as unknown. For this computation, the timeout was set to 10 hours. This activity required several weeks of computation. A result confirmed by at least two solvers was obtained for 75\% of the unknown benchmarks, with another 4\% having a tentative result from just one solver.
The results are shown in Table~\ref{Fig:post}.

>>>>

In the succeeding sections we describe 
the competition goals (\S\ref{sec:goals}), 
the SMT-LIB language that is the basis for the competition (\S\ref{sec:context}), 
the competition benchmarks (\S\ref{sec:benchmarks}), 
participants~(\S\ref{sec:participants}), 
procedure (\S\ref{sec:procedure}),
computational infrastructure (\S\ref{sec:starexec}), 
comparisons with other competitions~(\S\ref{sec:OtherCompetitions}), 
the results (\S\ref{sec:results}), 
the place of SMT-COMP in the FLoC Olympic Games (\S\ref{sec:floc}), 
and post-competition activities (\S\ref{sec:post}). 
Section \ref{sec:conclusions} presents observations on this competition and recommendations for the future.

@@

The SMT Competition is a competition among SMT solvers on a set of benchmark logic problems. Each benchmark problem is a combination of definitions and logical assertions expressed with respect to an underlying logical theory and, perhaps, some constraints on the kinds of expressions in that theory. For example,  the logic of linear arithmetic includes the multiplication operation, but only allows multiplication by constants. Each problem is a set of closed formulas (possibly including quantification over variables) over a set of constant or function symbols; a solution to the problem is an assignment of each constant and function evaluation to values in a way that satisfies all the problem's assertions. That is, the task is to find a \emph{satisfying assignment} for the benchmark problem or to determine that there is no such assignment. Since the presence of quantified expressions introduces incompleteness, solvers may also produce a potential solution that may be marked as possibly spurious.

As stated, the goal of SMT-COMP is similar to the goal of the SAT competition. The SMT logic extends SAT by incorporating defined theories, such as the theory of arrays, or of uninterpreted functions, or of arithmetic, or of bit-vectors. In addition, there may be constraints on the set of expressions allowed, such as only linear arithmetic, or only integer difference arithmetic. The theories also define \emph{sorts}, which make the theories a typed first-order logic. Examples of sorts used in current theories are Boolean, Int, Real, bit-vectors of various lengths, and arrays with arbitrary sorts as index and value. Each combination of underlying theories and language constraints is a \emph{logic}.
The names of the logics as used in SMT-LIB are combinations of initials. For example, AUFLIA is the logic with a
combination of Arrays (A), Uninterpreted Functions (UF), and linear integer arithmetic (LIA). Table \ref{logicAbbreviations} can be used to interpret these names.
\begin{table}
\begin{center}
\begin{tabular}{|l|l|}
\hline
QF\_ & Quantifier-Free \\
A & Arrays \\
UF & Uninterpreted Functions \\
BV & Bit-Vector \\
L & Linear (arithmetic) \\
N & Non-Linear (arithmetic) \\
IA/RA/IRA & Integer/Real/Mixed arithmetic \\
IDL/RDL & Integer/Real Difference Logic \\
\hline
\end{tabular}
\end{center}
\caption{Abbreviations used in logic names}
\label{logicAbbreviations}
\end{table}

@@

\begin{minipage}{\textwidth}
\begin{quote}
\lstset{frameround=fttt}
\begin{lstlisting}[frame=trBL]
(set-logic UFLIA)
(declare-fun max (Int Int) Int)
(assert (forall ((x Int) (y Int))
   (let ((m (max x y)))
      (and (>= m x) (>= m y) (or (= m x) (= m y))))))
(assert (not (forall ((x Int) (y Int))
   (let ((m (max x y)))
      (= (max m x) m)))))
(check-sat)
\end{lstlisting}
\end{quote}
\end{minipage}

@@

\begin{figure}
\begin{center}
\begin{tabular}{cc}
\includegraphics[width=.47\textwidth]{smtlib2-file-count-by-size.png}
&
\includegraphics[width=.47\textwidth]{smtlib2-app-file-count-by-size.png}
\\
\\
Non-incremental benchmarks & Incremental benchmarks
\end{tabular}
\end{center}
\caption{Distribution of SMT-LIB benchmarks by file size.
\label{fig:smtlib-bysize}}
\end{figure}

@@

The numbers of unknown, trivial, and remaining eligible benchmarks are shown in Table~\ref{Table:benchmarks}.  Note that these numbers vary widely from division to division.  As we discuss in Section~\ref{sec:benchmark-selection} below, all eligible benchmarks were actually used in the competition; these numbers are shown in bold.

\begin{table}
\caption{Numbers of main-track benchmarks. The second column shows the number of competitive solvers and, in square brackets, the number of demonstration-only solvers. Entries marked * exclude some benchmarks containing partial functions.}
\label{Table:benchmarks}
\centering
\begin{tabular}{|l|r|r|r|r|r|}
\hline
       & \multicolumn{1}{|c|}{\# of} & \multicolumn{4}{|c|}{\# of benchmarks} \\
 Logic & solvers & {\bf eligible} & unknown  & trivial & total  \\
\hline
ALIA & 3+[1] & {\bf 29} & 0 & 13 & 42 \\
AUFLIA & 3+[1] & {\bf 4} & 0 & 0 & 4\\
AUFLIRA & 3+[1] & {\bf 10791} & 168 & 9055 & 20014 \\
AUFNIRA & 2+[2] & {\bf 564} & 468 & 463 & 1495 \\
BV & 2+[1] & {\bf 0} & 191 & 0 & 191 \\
LIA & 3+[1] & {\bf 46} & 0 & 0 & 46 \\
LRA & 3+[1] & {\bf 171} & 450 & 0 & 621\\
NIA & 2+[1] & {\bf 9} & 0 & 0 & 9\\
NRA & 2+[1] & {\bf 3747} & 66 & 0 & 3813 \\
QF\_ABV & 7+[2] & {\bf 6457}* & 4191 & 4423 & 15091 \\
QF\_ALIA & 3+[2] & {\bf 97} & 0 & 29 & 126 \\
QF\_AUFBV & 2+[2] & {\bf 37} & 0 & 0 & 37 \\
QF\_AUFLIA & 4+[2] & {\bf 610} & 0 & 399 & 1009\\
QF\_AX & 3+[2] & {\bf 335} & 0 & 216 & 551 \\
QF\_BV & 8+[3] & {\bf 2488}* & 28138 & 546 & 32500 \\
QF\_IDL & 3+[1] & {\bf 1315} & 537 & 337 & 2189 \\
QF\_LIA & 4+[3] & {\bf 4381} & 1279 & 481 & 6141 \\
QF\_LRA & 4+[2] & {\bf 1343} & 208 & 131 & 1682\\
QF\_NIA & 3+[1] & {\bf 8327} & 927 & 105 & 9359 \\
QF\_NRA & 3+[1] & {\bf 10121} & 1392 & 27 & 11540 \\
QF\_RDL & 3+[1] & {\bf 132} & 85 & 38 & 255 \\
QF\_UF & 5+[2] & {\bf 4124} & 4 & 2522 & 6650 \\
QF\_UFBV & 2+[2] & {\bf 31} & 0 & 0 & 31 \\
QF\_UFIDL & 3+[1] & {\bf 311} & 0 & 130 & 441\\
QF\_UFLIA & 4+[2] & {\bf 484} & 0 & 114 & 598 \\
QF\_UFLRA & 4+[2] & {\bf 1176} & 87 & 367 & 1630 \\
QF\_UFNIA & 2+[1] & {\bf 7} & 0 &0 & 7 \\
QF\_UFNRA & 2+[1] & {\bf 32} & 11 & 0 & 43\\
UF & 3+[1] & {\bf 2830} & 2911 & 7 & 5748 \\
UFBV & 2+[1] & {\bf 0} & 191 & 0 & 191 \\
UFIDL & 2+[1] & {\bf 49} & 12 & 19 & 80 \\
UFLIA & 3+[1] & {\bf 5766} & 5499 & 873 & 12138 \\
UFLRA & 3+[1] & {\bf 25} & 0 & 0 & 25\\
UFNIA & 2+[1] & {\bf 1587} & 1052 & 712 & 3351 \\
\hline
Total & 18+[3] & {\bf 67426} & 47867 & 21007 & 137648 \\
\hline
\end{tabular}
\end{table}

@@

The language includes commands that allow a fine-grained interaction
with the solver, whereby client tools may incrementally push and pop
symbol declarations and assertions while running various
satisfiability checks, inspecting models, or obtaining unsatisfiability
proofs. Benchmark problems that have more than one
\lstinline{check-sat} command are called \emph{application} or \emph{incremental
  benchmarks}. Not all SMT solvers support all of these interaction
facilities, and the main track of the competition does not use such
application benchmarks.

@@

Benchmarks are also labeled by \emph{category} (not shown in the tables): simple checks, randomly generated problems from some template (e.g., N-queens problems for various values of N), problems crafted to test a certain capability, and problems obtained from industrial applications. The selection rules in previous years favored industrial benchmarks.

@@

The scrambling had another effect, previously unappreciated. Although most of the solvers are deterministic for a given input, scrambling can change the order of search. For example, iterating over the contents of a hashed set may occur in a different order if identifiers have different names. Consequently, the success of the solver on a given benchmark, the time it takes to find an answer, and even whether a latent bug in the solver is triggered may all depend on just how the benchmark is scrambled. Thus the scrambling adds a measure of uncertainty to the competition and prevents full testing on the benchmark set before the competition. Indeed, one solver encountered a bug during the competition that was directly related to the specific, randomly-chosen scrambling used during the competition. The effect of the possible variations in performance caused by scrambling is expected to be balanced by the large number of benchmarks employed in the competition to compare the solvers, as also observed in the SAT community~\cite{leberre+03}. Nevertheless, this point is worthy of future further investigation.

@@

\begin{table}
\caption{Fraction of benchmarks solved fastest (cpu-time) by each solver (top 3)}
\label{tab:speed}
\centering\begin{tabular}{|l|lll|}
\hline
Logic & Fraction solved fastest & & \\ \hline
ALIA & 51\% [Z3] & 48\% veriT  & \\
AUFLIA & 66\% [Z3] & 33\% CVC3  & \\
AUFLIRA & 60\% [Z3] & 37\% veriT & 1\% CVC3 \\
AUFNIRA & 72\% [Z3] & 23\% CVC3 & 2\% CVC4 \\
LIA & 43\% CVC4 & 30\% [Z3] & 17\% CVC3 \\
LRA & 42\% CVC4 & 38\% veriT & 18\% [Z3] \\
NIA & 66\% [Z3] & 33\% CVC4  & \\
NRA & 96\% CVC4 & 3\% [Z3]  & \\
QF\_ABV & 78\% Yices2 & 9\% SONOLAR & 7\% Kleaver-STP \\
QF\_ALIA & 95\% Yices2 & 4\% [Z3] & \\
QF\_AUFBV & 81\% Yices2 & 10\% [MathSAT] & 5\% [Z3] \\
QF\_AUFLIA & 60\% Yices2 & 38\% [Z3] & 1\% CVC4 \\
QF\_AX & 92\% Yices2 & 6\% [Z3] & 0\% CVC4 \\
QF\_BV & 33\% Yices2 & 25\% [Z3] & 13\% 4Simp \\
QF\_IDL & 87\% Yices2 & 11\% [Z3] & 0\% veriT \\
QF\_LIA & 76\% Yices2 & 11\% [Z3] & 7\% CVC4 \\
QF\_LRA & 86\% Yices2 & 3\% veriT & 3\% CVC4 \\
QF\_NIA & 90\% [Z3] & 7\% AProVE & 1\% CVC3 \\
QF\_NRA & 88\% [Z3] & 6\% CVC4 & 4\% CVC3 \\
QF\_RDL & 71\% Yices2 & 25\% [Z3] & 3\% veriT \\
QF\_UF & 96\% Yices2 & 1\% veriT & 1\% [Z3] \\
QF\_UFBV & 96\% Yices2 & 3\% [MathSAT]  & \\
QF\_UFIDL & 59\% [Z3] & 36\% Yices2 & 3\% veriT \\
QF\_UFLIA & 97\% Yices2 & 2\% [Z3] & 0\% CVC4 \\
QF\_UFLRA & 91\% Yices2 & 4\% [Z3] & 2\% veriT \\
QF\_UFNIA & 100\% CVC4 & & \\
QF\_UFNRA & 59\% CVC3 & 37\% [Z3] & 3\% CVC4 \\
UF & 37\% veriT & 32\% CVC4 & 24\% [Z3] \\
UFIDL & 57\% CVC4 & 42\% [Z3]  & \\
UFLIA & 56\% [Z3] & 35\% veriT & 7\% CVC4 \\
UFLRA & 95\% veriT & 4\% [Z3]  & \\
UFNIA & 80\% [Z3] & 10\% CVC3 & 9\% CVC4 \\

\hline
\end{tabular}
\end{table}

@@

\begin{table}
\caption{Contribution from each solver (top 3)}
\label{tab:cont}
\centering\begin{tabular}{|l|lll|}
\hline
Logic & Unique contribution & & \\
\hline
ALIA & 0.4253 [Z3] & 0.4080 CVC4 & 0.1667 veriT \\
AUFLIA & 0.3889 CVC4 & 0.2778 [Z3] & 0.2222 CVC3 \\
AUFLIRA & 0.2615 [Z3] & 0.2521 CVC4 & 0.2503 CVC3 \\
AUFNIRA & 0.2509 [Z3] & 0.2506 CVC4 & 0.2497 CVC3 \\
LIA & 0.3949 [Z3] & 0.3514 CVC4 & 0.2319 CVC3 \\
LRA & 0.3070 CVC4 & 0.3070 [Z3] & 0.2895 CVC3 \\
NIA & 0.8333 [Z3] & 0.1667 CVC4 & 0.0000 Yices2 \\
NRA & 0.5071 [Z3] & 0.4929 CVC4 & 0.0000 Yices2 \\
QF\_ABV & 0.1147 Yices2 & 0.1146 Boolector-just... & 0.1144 Boolector-dual... \\
QF\_ALIA & 0.2235 Yices2 & 0.2235 SMTInterpol & 0.2235 [Z3] \\
QF\_AUFBV & 0.2635 [MathSAT] & 0.2635 CVC4 & 0.2500 Yices2 \\
QF\_AUFLIA & 0.1996 Yices2 & 0.1996 [MathSAT] & 0.1996 CVC4 \\
QF\_AX & 0.2000 Yices2 & 0.2000 [MathSAT] & 0.2000 CVC4 \\
QF\_BV & 0.1051 Boolector & 0.0976 STP-Crypto... & 0.0934 abziz\_min\_f... \\
QF\_IDL & 0.2893 [Z3] & 0.2731 Yices2 & 0.2395 CVC4 \\
QF\_LIA & 0.1620 [CVC4-...fix] & 0.1610 [Z3] & 0.1610 SMTInterpol \\
QF\_LRA & 0.1772 CVC4 & 0.1689 Yices2 & 0.1678 SMTInterpol \\
QF\_NIA & 0.4990 [Z3] & 0.4912 AProVE & 0.0084 CVC3 \\
QF\_NRA & 0.7714 [Z3] & 0.1347 CVC3 & 0.0898 CVC4 \\
QF\_RDL & 0.2734 Yices2 & 0.2431 veriT & 0.2431 [Z3] \\
QF\_UF & 0.1451 Yices2 & 0.1449 veriT & 0.1446 CVC4 \\
QF\_UFBV & 0.4140 Yices2 & 0.3172 [Z3] & 0.1720 [MathSAT] \\
QF\_UFIDL & 0.2708 Yices2 & 0.2708 [Z3] & 0.2309 CVC4 \\
QF\_UFLIA & 0.1698 Yices2 & 0.1698 [MathSAT] & 0.1698 CVC4 \\
QF\_UFLRA & 0.1715 [Z3] & 0.1681 Yices2 & 0.1681 [MathSAT] \\
QF\_UFNIA & 0.5000 CVC4 & 0.5000 [Z3] & 0.0000 Yices2 \\
QF\_UFNRA & 0.5802 [Z3] & 0.2099 CVC4 & 0.2099 CVC3 \\
UF & 0.4921 CVC4 & 0.1988 [Z3] & 0.1759 CVC3 \\
UFIDL & 0.5204 [Z3] & 0.4796 CVC4 & 0.0000 Yices2 \\
UFLIA & 0.4069 CVC4 & 0.3502 [Z3] & 0.2425 veriT \\
UFLRA & 0.2857 [Z3] & 0.2381 CVC4 & 0.2381 veriT \\
UFNIA & 0.4156 [Z3] & 0.3860 CVC4 & 0.1984 CVC3 \\
\hline
\end{tabular}
\end{table}

@@

\begin{figure}
\begin{center}
\includegraphics[width=.95\textwidth]{QF_BV-timeout-logscale2.png}
\end{center}
\caption{Solver success for QF\_BV benchmarks over time.}
\label{Fig:timeouts}
\end{figure}

@@

The data in Table~\ref{Fig:winners} also shows that most solvers have a similar trend in success as the timeout is increased. For example, most solvers solve about 50\% (ranging from 40\% to~62\%) of the benchmarks that solver solves within the timeout by 1.0 second and
about 75\% (ranging from 65\% to 79\%) of those solved within the timeout by 10 seconds. However, some solvers, such as Boolector, are outliers: Boolector is slow off the mark---solving only~33\% and 61\% of the benchmarks it eventually solves
by 1 second and 10 seconds, respectively, even though by 2400 seconds it solves more benchmarks than any other tool, in this division.

@@

After the competition, David Cok (competition chair) and Clark Barrett and Morgan Deters (SMT-LIB coordinators) collaborated, with the assistance of Aaron Stump (StarExec lead), in attempting to discover the status of the SMT-LIB benchmarks that were marked as unknown. For this computation, the timeout was set to 10 hours. This activity required several weeks of computation. A result confirmed by at least two solvers was obtained for~75\% of the unknown benchmarks, with another 4\% having a tentative result from just one solver.
The results are shown in Table~\ref{Fig:post}.
